---
title: "improving_sensitivity"
output: html_document
---

We would like to improve the sensitivity of the model from the currently very low sensitivity value of .431

```{r}
source("model_test_v0.R") # Load the base model version

my_metrics(results, truth = patch, estimate = .pred_class, .pred_TRUE, event_level = "second")
```

## Adjust Classification Threshold

By default, ≥ .5 is a patch. Lowering this threshold can improve the sensitivity of the model, at the expense of specificity.

```{r}
# Roc curve illustrates tradeoff between sensitivity and specificity
roc_curve(results, truth = patch, .pred_FALSE) |> autoplot()

# using phi threshold from Johnson et al. 2025
altered_patchpreds <- results |>
  mutate(newpredclass = factor(.pred_TRUE >= .2))

my_metrics(altered_patchpreds, truth = patch, estimate = newpredclass, .pred_TRUE, event_level = "second")
#conf_mat(altered_patchpreds, patch, newpredclass)
```
## Optimizing tuning for alternative performance metric

Optimizing tuning to maximize sensitivity or F1 score rather than AUC

## Handle class imbalance via modifications to input dataset

Upsample the minority class or downsample the majority class with `step_upsample()` and `step_downsample()`

Can also doing this using SMOTE (`themis`) package in tidymodels, class weights (`scale_pos_weight`)

## Feature Engineering

New interaction terms, domain-informed variables, etc.

Use feature importance to assess what drives decisions.

## Post-processing ensemble

Try combining model with mechanistic model or one prioritizing sensitivity


