---
title: "Hmmmm"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
suppressPackageStartupMessages({
  library(stars)
  library(readr)
})
```

```{r}
infile = "/mnt/ecocast/projects/students/ojohnson/daily-forecasts/model_exploration/sample_stars.rds"
x = readr::read_rds(infile)
nodate = st_as_stars()
nodate$valuse2 = 1
```

```{r}
suppressWarnings({
  y = stars::write_mdim(x, "out.nc")
  nodate_y = stars::write_mdim(nodate, "outnodate.nc")
})
```

```{r}
suppressWarnings({
  stars = stars::read_stars("out.nc")
  mdim = stars::read_mdim("out.nc")
  ncdf = stars::read_ncdf("out.nc")
  
  mdimnodate = stars::read_mdim("outnodate.nc")
})
```

```{r}
st_dimensions(x)
st_dimensions(y)
st_dimensions(stars)
st_dimensions(mdim)
st_dimensions(ncdf)

st_dimensions(mdimnodate)
```

```{r}
d = st_dimensions(stars)
d[[3]]$offset = d[[3]]$offset |>
  units::drop_units() |>
  as.Date(origin = as.Date("1970-01-01"))
d[[3]]$delta = 1
d[[3]]$refsys = "Date"
st_dimensions(stars) <- d
```

```{r}
st_dimensions(x)
st_dimensions(stars)
```

# Digging into the source code

Helpers from [stars github](https://github.com/r-spatial/stars/blob/main/R/mdim.R)
```{r}
# convert stars object into a list of CDL-like variables with named dimensions
is_curvilinear <- function(x) {return(FALSE)}
add_attr = function(x, at) { # append at to attribute "attrs"
		structure(x, attrs = c(attr(x, "attrs"), at))
}
add_units_attr = function(l) {
		f = function(x) {
			if (inherits(x, "units"))
				add_attr(x, c(units = as.character(units(x))))
			else if (inherits(x, c("POSIXct", "PCICt"))) {
				cal = if (!is.null(cal <- attr(x, "cal")))
					c(calendar = paste0(cal, "_day")) # else NULL, intended
				x = as.numeric(x)
				if (all(x %% 86400 == 0))
					add_attr(x/86400, c(units = "days since 1970-01-01", cal))
				else if (all(x %% 3600 == 0))
					add_attr(x / 3600, c(units = "hours since 1970-01-01 00:00:00", cal))
				else
					add_attr(x, c(units = "seconds since 1970-01-01 00:00:00", cal))
			} else if (inherits(x, "Date"))
				add_attr(as.numeric(x), c(units = "days since 1970-01-01"))
			else
				x
		}
		lapply(l, f)
}
st_as_cdl = function(x) {
	e = add_units_attr(expand_dimensions(x)) # TODO: handle sfc dimensions
	d = st_dimensions(x)
	dimx = dim(x)
	xy = attr(st_dimensions(x), "raster")$dimensions
	co = NULL
	for (i in seq_along(e)) {
		if (is.null(dim(e[[i]])) && !is.list(e[[i]]))
			e[[i]] = structure(e[[i]], dim = setNames(length(e[[i]]), names(e)[i]))
		else if (is_curvilinear(x)) { # curvilinear coordinate matrices:
			if (names(e)[i] == xy[1]) names(e)[i] = "lon"
			if (names(e)[i] == xy[2]) names(e)[i] = "lat"
		} else if (inherits(e[[i]], "sfc")) { # vector data cube:
			sfc = e[[i]]
			e[[i]] = structure(seq_along(sfc), dim = setNames(length(sfc), names(e)[i]))
			e = cdl_add_geometry(e, i, sfc)
			if (inherits(sfc, "sfc_POINT")) {
				if (isTRUE(st_is_longlat(x))) {
					co = c(coordinates = "lat lon")
					xy = c("lon", "lat") # hack
				} else {
					co = c(coordinates = "x y")
					xy = c("x", "y") # hack
				}
			} else {
				dimx = c(dimx, attr(e, "dims"))
				co = c(coordinates = "x y")
				xy = c("x", "y") # hack
			} 
			co = c(co, geometry = "geometry")
		}
	}

	if (is_curvilinear(x)) {
		e[["lat"]] = add_attr(e[["lat"]], c(units = "degrees_north", "_CoordinateAxisType" = "Lat", axis = "Y"))
		e[["lon"]] = add_attr(e[["lon"]], c(units = "degrees_east",  "_CoordinateAxisType" = "Lon", axis = "X"))
		co = paste(rev(setdiff(names(d), c("x", "y", "lat", "lon"))), "lat lon")
	} else {
		if (!any(is.na(xy))) {
			e[[ xy[1] ]] = add_attr(e[[ xy[1] ]], c(axis = "X"))
			e[[ xy[2] ]] = add_attr(e[[ xy[2] ]], c(axis = "Y"))
		}
	}
	for (i in seq_along(x))
		x[[i]] = add_attr(x[[i]], co)

	x = add_units_attr(x) # unclasses x
	for (i in seq_along(x))
		if (is.null(names(dim(x[[i]])))) # FIXME: read_ncdf() doesn't name array dim
			names(dim(x[[i]])) = names(d)

	for (i in names(e)) # copy over dimension values
		x[[i]] = e[[i]]

	which_dims = function(a, dimx) {
		m = match(names(dim(a)), names(dimx), nomatch = numeric(0)) - 1
		if (all(is.na(m)))
			integer(0)
		else
			m
	}
	ret = lapply(x, function(a) structure(a, which_dims = which_dims(a, dimx)))
	structure(ret, which_crs = !(names(ret) %in% names(e)), 
			  is_numeric = sapply(ret, is.numeric),
			  dims = attr(e, "dims"))
}
cdl_add_geometry = function(e, i, sfc) {
	stopifnot(inherits(sfc, c("sfc_POINT", "sfc_POLYGON", "sfc_MULTIPOLYGON", "sfc_LINESTRING", "sfc_MULTILINESTRING")))

	if (inherits(sfc, "sfc_POLYGON"))
		sfc = st_cast(sfc, "MULTIPOLYGON")
	if (inherits(sfc, "sfc_LINESTRING"))
		sfc = st_cast(sfc, "MULTILINESTRING")

	cc = st_coordinates(sfc)
	if (inherits(sfc, c("sfc_MULTIPOLYGON"))) {
		e$node = structure(seq_len(nrow(cc)), dim = c(node = nrow(cc)))
		e$x = structure(cc[, 1], dim = c(node = nrow(cc)))
		e$y = structure(cc[, 2], dim = c(node = nrow(cc)))
		e$node_count = structure(rle(cc[,"L3"])$lengths, dim = setNames(length(sfc), names(e)[i]))
		part = rle(cc[,"L3"] * 2 * max(cc[,"L2"]) + 2 * (cc[,"L2"] - 1) + cc[,"L1"])$lengths
		e$part_node_count = structure(part, dim = c(part = length(part)))
		e$interior_ring = structure(as.numeric(cc[cumsum(part), "L1"] > 1), dim = c(part = length(part)))
		attr(e, "dims") = c(node = nrow(cc), part = length(part))
		e$geometry = add_attr(structure(numeric(0), dim = c("somethingNonEx%isting" = 0)),
			c(geometry_type = "polygon", node_count = "node_count", node_coordinates = "x y",
		  	part_node_count = "part_node_count", interior_ring = "interior_ring",
		  	grid_mapping = if (!is.na(st_crs(sfc))) "crs" else NULL))
	} else if (inherits(sfc, "sfc_MULTILINESTRING")) { # LINE:
		e$node = structure(seq_len(nrow(cc)), dim = c(node = nrow(cc)))
		e$x = structure(cc[, 1], dim = c(node = nrow(cc)))
		e$y = structure(cc[, 2], dim = c(node = nrow(cc)))
		e$node_count = structure(rle(cc[,"L2"])$lengths, dim = setNames(length(sfc), names(e)[i]))
		part = rle((cc[,"L2"] - 1) * max(cc[,"L1"]) + cc[,"L1"])$lengths
		e$part_node_count = structure(part, dim = c(part = length(part)))
		attr(e, "dims") = c(node = nrow(cc), part = length(part))
		e$geometry = add_attr(structure(numeric(0), dim = c("somethingNonEx%isting" = 0)),
			c(geometry_type = "line", node_count = "node_count", node_coordinates = "x y",
		  	part_node_count = "part_node_count",
		  	grid_mapping = if (!is.na(st_crs(sfc))) "crs" else NULL))
	} else { # POINT:
		if (ll <- isTRUE(st_is_longlat(sfc))) {
			e$lon = add_attr(structure(cc[, 1], dim = setNames(nrow(cc), names(e)[i])), 
						 	c(units = "degrees_north", standard_name = "longitude"))
			e$lat = add_attr(structure(cc[, 2], dim = setNames(nrow(cc), names(e)[i])), 
						 	c(units = "degrees_east", standard_name = "latitude"))
		} else {
			e$x = structure(cc[, 1], dim = setNames(nrow(cc), names(e)[i]))
			e$y = structure(cc[, 2], dim = setNames(nrow(cc), names(e)[i]))
		}
		e$geometry = add_attr(structure(numeric(0), dim = c("somethingNonEx%isting" = 0)),
				c(geometry_type = "point", node_coordinates = ifelse(ll, "lon lat", "x y"),
				grid_mapping = if (!is.na(st_crs(sfc))) "crs" else NULL))
	}
	e
}
```

Example code snippet from [write_mdim reference](https://r-spatial.github.io/stars/reference/mdim.html)
```{r}
set.seed(135)
m = matrix(runif(10), 2, 5)
names(dim(m)) = c("stations", "time")
times = as.Date("2022-05-01") + 1:5
pts = st_as_sfc(c("POINT(0 1)", "POINT(3 5)"))
s = st_as_stars(list(Precipitation = m)) |>
 st_set_dimensions(1, values = pts) |>
 st_set_dimensions(2, values = times)
nc = tempfile(fileext=".nc")
write_mdim(s, nc)

s_read <- read_mdim(nc)
st_dimensions(s_read)
s_read 
```
How does cdl process the stars object
```{r}
cdl_x <- st_as_cdl(x)

cdl_s <- st_as_cdl(s)

cdl_nodate <- st_as_cdl(nodate)

cdl_x$date
cdl_s$time

cdl_x$lon
cdl_s$x

cdl_nodate$x
```

Actual code for read_mdim()
```{r}
#' Read or write data using GDAL's multidimensional array API
#'
#' Read or write data using GDAL's multidimensional array API
#' @name mdim
#' @param filename name of the source or destination file or data source
#' @param variable name of the array to be read; if `"?"`, a list of array names is returned, with group name as list element names.
#' @param options character; driver specific options regarding the opening (read_mdim) or creation (write_mdim) of the dataset
#' @param raster names of the raster variables (default: first two dimensions)
#' @param offset integer; zero-based offset for each dimension (pixels) of sub-array to read, defaults to 0 for each dimension(requires sf >= 1.0-9)
#' @param count integer; size for each dimension (pixels) of sub-array to read (default: read all); a value of NA will read the corresponding dimension entirely; counts are relative to the step size (requires sf >= 1.0-9)
#' @param step integer; step size for each dimension (pixels) of sub-array to read; defaults to 1 for each dimension (requires sf >= 1.0-9)
#' @param proxy logical; return proxy object?
#' @param debug logical; print debug info?
#' @param bounds logical or character: if \code{TRUE} tries to infer from "bounds" attribute; if character, 
#' named vector of the form \code{c(longitude="lon_bnds", latitude="lat_bnds")} with names dimension names
#' @param curvilinear control reading curvilinear (geolocation) coordinate arrays; if \code{NA} try reading the x/y dimension names; if character, defines the arrays to read; if \code{FALSE} do not try; see also \link{read_stars}
#' @param normalize_path logical; if \code{FALSE}, suppress a call to \link{normalizePath} on \code{filename}
#' @details it is assumed that the first two dimensions are easting and northing
#' @param ... ignored
#' @seealso \link[sf]{gdal_utils}, in particular util \code{mdiminfo} to query properties of a file or data source containing arrays
#' @export
read_mdim = function(filename, variable = character(0), ..., options = character(0), 
					 raster = NULL, offset = integer(0), count = integer(0), step = integer(0), proxy = FALSE, 
					 debug = FALSE, bounds = TRUE, curvilinear = NA, normalize_path = TRUE) {

	stopifnot(is.character(filename), is.character(variable), is.character(options))
	if (normalize_path)
		filename = enc2utf8char(maybe_normalizePath(filename, np = normalize_path))
	ret = gdal_read_mdim(filename, variable, options, rev(offset), rev(count), rev(step), proxy, debug)

	if (identical(variable, "?"))
		return(ret) # RETURNS

	if (length(ret$dimensions) == 1 && length(ret$array_list) == 1 && is.data.frame(ret$array_list[[1]]))
		return(ret$array_list[[1]]) ## composite data: RETURNS

	ret = recreate_geometry(ret)
	if (isTRUE(bounds) || is.character(bounds))
		ret$dimensions = mdim_use_bounds(ret$dimensions, filename, bounds)

	create_units = function(x) {
		u = attr(x, "units")
		x = structure(x, units = NULL) # remove attribute
		if (is.null(u) || u %in% c("", "none"))
			x
		else {
			if (!is.null(a <- attr(x, "attributes")) && !is.na(cal <- a["calendar"]) && 
						cal %in% c("360_day", "365_day", "noleap"))
				get_pcict(x, u, cal)
			else {
				days_since = grepl("days since", u)
				u = try_as_units(u)
				if (!inherits(u, "units")) # FAIL:
					x
				else {
					units(x) = u
					if (days_since && inherits(d <- try(as.Date(x), silent = TRUE), "Date")) 
						d
					else if (inherits(p <- try(as.POSIXct(x), silent = TRUE), "POSIXct"))
						p
					else
						x
				}
			}
		}
	}
	l = rev(lapply(ret$dimensions, function(x) {
			   if (inherits(x, "sfc")) x else create_units(x$values[[1]])
			}))
	if (length(offset) != 0 || length(step) != 0 || length(count) != 0) {
		if (length(offset) == 0)
			offset = rep(0, length(l))
		if (length(step) == 0)
			step = rep(1, length(l))
		ll = lengths(lapply(l, as.list)) # take care of dimensions of class intervals
		if (length(count) == 0)
			count = floor((ll - offset)/step)
		else if (any(a <- is.na(count)))
			count[a] = floor((ll[a] - offset[a])/step[a])
		for (i in seq_along(l)) {
			l[[i]] = l[[i]][seq(from = offset[i]+1, length.out = count[i], by = step[i])]
		}
	}

	# create dimensions table:
	sf = any(sapply(l, function(x) inherits(x, "sfc")))
	# FIXME: i %in% 1:2 always the case?
	raster_dims = match_raster_dims(names(l))
	d = mapply(function(x, i) create_dimension(values = x, is_raster = !sf && i %in% raster_dims,
									   point = ifelse(length(x) == 1, TRUE, NA)),
			   l, seq_along(l), SIMPLIFY = FALSE)
	if (is.null(raster)) {
		raster = if (sf)
					get_raster(dimensions = rep(NA_character_,2))
				else
					get_raster(dimensions = names(d)[raster_dims])
	} else
		raster = get_raster(dimensions = raster)
	dimensions = create_dimensions(d, raster = raster)

	# handle array units:
	for (i in seq_along(ret$array_list))
		if (nchar(u <- attr(ret$array_list[[i]], "units")) && inherits(u <- try_as_units(u), "units"))
			units(ret$array_list[[i]]) = u
	clean_units = function(x) { 
		if (identical(attr(x, "units"), "")) 
			structure(x, units = NULL)
		else
			x
	}

	# create return object:
	if (proxy) {
		lst = lapply(ret$array_list, function(x) filename)
		st = st_stars_proxy(lst, dimensions, NA_value = NA_real_, resolutions = NULL, class = "mdim")
		if (!is.null(ret$srs))
			st_set_crs(st, ret$srs)
		else
			st
	} else { 
		lst = lapply(ret$array_list, function(x) structure(clean_units(x), dim = rev(dim(x))))
		st = st_stars(lst, dimensions)
		if (is.null(ret$srs) || is.na(ret$srs)) {
			if (missing(curvilinear) || is.character(curvilinear)) { # try curvilinear:
				xy = raster$dimensions
				ll = curvilinear
				if (is.character(curvilinear)) {
					if (is.null(names(curvilinear)))
						names(curvilinear) = xy[1:2]
					ret = try(st_as_stars(st, curvilinear = curvilinear), silent = TRUE)
					if (inherits(ret, "stars"))
						st = ret
				}
				if (!is_curvilinear(st) &&
					inherits(x <- try(read_mdim(filename, ll[1], curvilinear = FALSE), silent = TRUE), "stars") &&
					inherits(y <- try(read_mdim(filename, ll[2], curvilinear = FALSE), silent = TRUE), "stars") &&
						identical(dim(x)[xy], dim(st)[xy]) && identical(dim(y)[xy], dim(st)[xy]))
					st = st_as_stars(st, curvilinear = setNames(list(x, y), xy))
			}
			st
		} else
			st_set_crs(st, ret$srs)
	}
}

try_as_units = function(u) {
	un = try(suppressWarnings(as_units(u)), silent = TRUE)
	if (inherits(un, "try-error")) # try without ^:
		un = try(suppressWarnings(as_units(gsub("^", "", u, fixed = TRUE))), silent = TRUE)
	if (inherits(un, "try-error")) # try without **
		un = try(suppressWarnings(as_units(gsub("**", "", u, fixed = TRUE))), silent = TRUE)
	if (inherits(un, "try-error") && u != "none") {
		warning(paste("ignoring unrecognized unit:", u), call. = FALSE)
		NULL
	} else
		un
}

create_units = function(x) {
		u = attr(x, "units")
		x = structure(x, units = NULL) # remove attribute
		if (is.null(u) || u %in% c("", "none"))
			x
		else {
			if (!is.null(a <- attr(x, "attributes")) && !is.na(cal <- a["calendar"]) && 
						cal %in% c("360_day", "365_day", "noleap"))
				get_pcict(x, u, cal)
			else {
				days_since = grepl("days since", u)
				u = try_as_units(u)
				if (!inherits(u, "units")) # FAIL:
					x
				else {
					units(x) = u
					if (days_since && inherits(d <- try(as.Date(x), silent = TRUE), "Date")) 
						d
					else if (inherits(p <- try(as.POSIXct(x), silent = TRUE), "POSIXct"))
						p
					else
						x
				}
			}
		}
}

create_dimension = function(from = 1, to, offset = NA_real_, delta = NA_real_, 
		refsys = NA_character_, point = NA, values = NULL, is_raster = FALSE)  {

	example = NA

	if (! is.null(values)) { # figure out from values whether we have sth regular:
		from = 1
		to = length(values)
		if (!(is.character(values) || is.factor(values)) && is.atomic(values)) { 
			if (! all(is.finite(values)))
				warning("dimension value(s) non-finite")
			else {
				if (regular_intervals(values)) {
					offset = values[1]
					if (length(values) > 1) {
						delta = diff(values[1:2])
					# shift half grid cell size if x or y raster dim cell midpoints:
						if (is_raster)
							offset = offset - 0.5 * delta
					}
					values = NULL
					example = offset
				} else
					example = values
			}
		}
		if (inherits(values, "intervals"))
			example = values$start

		# refsys:
		if (inherits(example, "POSIXct"))
			refsys = "POSIXct"
		else if (inherits(example, "Date"))
			refsys = "Date"
		else if (inherits(example, "PCICt"))
			refsys = paste0("PCICt_", attr(example, "cal"))
		else if (inherits(example, "units"))
			refsys = "udunits"

		if (inherits(values, "sfc")) {
			point = inherits(values, c("sfc_POINT", "sfc_MULTIPOINT"))
			if (!is.na(st_crs(values)) && is.na(refsys)) # inherit:
				refsys = st_crs(values)
		}
		if (is.numeric(values) && (is.na(point) || !point)) {
			values = if (is_raster)
					set_dimension_values(centers = values)
				else
					set_dimension_values(start = values)
		}
	}
	structure(list(from = unname(from), to = unname(to), offset = unname(offset), 
		delta = unname(delta), refsys = refsys, point = unname(point), values = values), 
		class = "dimension")
}

regular_intervals = function(x, epsilon = 1e-10) {
	if (length(x) <= 1)
		FALSE
	else {
		ud = if (is.atomic(x) && (is.numeric(x) || inherits(x, c("POSIXt", "Date", "PCICt"))))
				unique(diff(x))
			else {
				if (inherits(x, "intervals") && identical(tail(x$end, -1), head(x$start, -1)))
					x$end - x$start
				else
					return(FALSE)
			}
		ud = as.numeric(ud)
		isTRUE((abs(diff(range(ud)) / mean(ud))) < epsilon)
	}
}
set_dimension_values = function(..., start = NULL, end = NULL, centers = NULL) {
	if (!is.null(centers) && (!is.null(start) || !is.null(end)))
		stop("if centers is given, start and end should not be given")
	if (!is.null(centers)) {
		l = length(centers)
		if (l <= 1)
			stop("cannot derive cell boundaries from a single center: specify start and end")
		d = diff(centers)
		start = c(centers - 0.5 * c(d[1], d))
		end = c(start[-1], centers[l] + 0.5 * d[l-1])
	} else if (missing(end)) {
		l = length(start)
		end = if (l == 1)
				NA_real_ #stop("cannot infer end values from a single start value")
			else
				c(start[-1], start[l] + diff(c(start[l-1], start[l])))
	}
	make_intervals(start = start, end = end)
}
```

Bootlegged version to diagnose
```{r}
filename = "out.nc"
filename2 = "outnodate.nc"
variable = character(0)
options = character(0)
raster = NULL
offset = integer(0)
count = integer(0)
step = integer(0)
proxy = FALSE
debug = FALSE
bounds = TRUE

ret = gdal_read_mdim(filename, variable, options, rev(offset), rev(count), rev(step), proxy, debug)
ret2 = gdal_read_mdim(filename2, variable, options, rev(offset), rev(count), rev(step), proxy, debug)

l = rev(lapply(ret$dimensions, function(x) {
			   if (inherits(x, "sfc")) x else create_units(x$values[[1]])
		  }))
l2 = rev(lapply(ret2$dimensions, function(x) {
			   if (inherits(x, "sfc")) x else create_units(x$values[[1]])
		  }))

### Possibly important
if (length(offset) != 0 || length(step) != 0 || length(count) != 0) {
	if (length(offset) == 0)
		offset = rep(0, length(l))
	if (length(step) == 0)
		step = rep(1, length(l))
	ll = lengths(lapply(l, as.list)) # take care of dimensions of class intervals
	if (length(count) == 0)
		count = floor((ll - offset)/step)
	else if (any(a <- is.na(count)))
		count[a] = floor((ll[a] - offset[a])/step[a])
	for (i in seq_along(l)) {
		l[[i]] = l[[i]][seq(from = offset[i]+1, length.out = count[i], by = step[i])]
	}
}

# create dimensions table GOES WRONG HERE.
raster_dims = 1:2
d = mapply(function(x, i) create_dimension(values = x, is_raster = i %in% raster_dims,
								   point = ifelse(length(x) == 1, TRUE, NA)),
		   l, seq_along(l), SIMPLIFY = FALSE)
d2 = mapply(function(x, i) create_dimension(values = x, is_raster = i %in% raster_dims,
								   point = ifelse(length(x) == 1, TRUE, NA)),
		   l2, seq_along(l2), SIMPLIFY = FALSE)

d$lon

d2$x
```

Dig into create_dimension
```{r}
from = 1
values = l[[1]]

create_dimension = function(from = 1, to, offset = NA_real_, delta = NA_real_, 
		refsys = NA_character_, point = NA, values = NULL, is_raster = FALSE)  {

	example = NA

	if (! is.null(values)) { # figure out from values whether we have sth regular:
		from = 1
		to = length(values)
		if (!(is.character(values) || is.factor(values)) && is.atomic(values)) { 
			if (! all(is.finite(values)))
				warning("dimension value(s) non-finite")
			else {
				if (regular_intervals(values)) {
					offset = values[1]
					if (length(values) > 1) {
						delta = diff(values[1:2])
					# shift half grid cell size if x or y raster dim cell midpoints:
						if (is_raster)
							offset = offset - 0.5 * delta
					}
					values = NULL
					example = offset
				} else
					example = values
			}
		}
		if (inherits(values, "intervals"))
			example = values$start

		# refsys:
		if (inherits(example, "POSIXct"))
			refsys = "POSIXct"
		else if (inherits(example, "Date"))
			refsys = "Date"
		else if (inherits(example, "PCICt"))
			refsys = paste0("PCICt_", attr(example, "cal"))
		else if (inherits(example, "units"))
			refsys = "udunits"

		if (inherits(values, "sfc")) {
			point = inherits(values, c("sfc_POINT", "sfc_MULTIPOINT"))
			if (!is.na(st_crs(values)) && is.na(refsys)) # inherit:
				refsys = st_crs(values)
		}
		if (is.numeric(values) && (is.na(point) || !point)) {
			values = if (is_raster)
					set_dimension_values(centers = values)
				else
					set_dimension_values(start = values)
		}
	}
	structure(list(from = unname(from), to = unname(to), offset = unname(offset), 
		delta = unname(delta), refsys = refsys, point = unname(point), values = values), 
		class = "dimension")
}
```

```{r}
regular_intervals(l[[1]]) # FALSE

regular_intervals(l2[[1]]) # TRUE

riv = l[[1]]
ud <- unique(diff(riv)) 
isTRUE((abs(diff(range(ud)) / mean(ud))) < 1e-10)
ud_raw <- st_get_dimension_values(x, "lon") |> diff() |> unique()
isTRUE((abs(diff(range(ud_raw)) / mean(ud_raw))) < 1e-10)


```

SO the issue is that the values of the data are corrupted enough on save that the difference between variables stops being a clean 1/12 and regular_intervals() incorrectly parses those values as discrete. gdal_write_mdim is the problem 🥲

Solution -- specify additional variables to manually set the offset...
OR be like ben and write our own wrapper to reassign the dimensions of those variables. 

```{r}
suppressWarnings({
  write_mdim(original, "out_as_dbl.nc", as_float = FALSE)
  read_in <- read_mdim("out_as_dbl.nc")
})
st_dimensions(original)
st_dimensions(read_in)

```



